---
title: "EcoGeno 4/19/17 Picrust"
author: "R Kirsten Tyler"
date: "4/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Picrust!**
Uses extended ancestral-state reconstruction algorithm to ID a closely related microbe that has been sequenced (its full genome has been sequenced) to find function of microbes

Picrust gives you a closed-reference (the most well-understood OTUs...the OTUs that are known) OTU table with KEGG orthology terms

In terminal:
```{}
filter_otus_from_otu_table.py -i ~/16s_analysis/otu_table_mc2_w_tax_no_pynast_failures_no_chimeras_frequency_filtered.biom -o ~/16s_analysis/closed_otu_table.biom --negate_ids_to_exclude -e /usr/lib/python2.7/site-packages/qiime_default_reference/gg_13_8_otus/rep_set/97_otus.fasta 

```

Any word in front of the OTU name means it's not a 'referenced' OTU

```{}
#How many closed ref OTUs?

biom summarize-table -i closed_otu_table.biom

#normalize by picrust’s method

normalize_by_copy_number.py -i ~/16s_analysis/closed_otu_table.biom -o ~/16s_analysis/closed_otu_table_norm.biom

```

Now we'll predict metagenomes. We'll get both a biom file and a text file to view

```{}

#for text file:
predict_metagenomes.py -f -i ~/16s_analysis/closed_otu_table_norm.biom -o ~/16s_analysis/metagenome_predictions.txt -a nsti_per_sample.txt 

#for biom output file: 
predict_metagenomes.py -i ~/16s_analysis/closed_otu_table_norm.biom -o ~/16s_analysis/metagenome_predictions.biom -a nsti_per_sample.txt

```

KO terms are representative of all taxa in database that share a common ancestor with your OTU

```{}
#Collapsing to higher KO hierarchy terms to look at "pathways of functionality"

text file: 
categorize_by_function.py -f -i metagenome_predictions.biom -c KEGG_Pathways -l 3 -o metagenome_predictions.L3.txt

biom file:
categorize_by_function.py -i metagenome_predictions.biom -c KEGG_Pathways -l 3 -o metagenome_predictions.L3.biom

```

Now let's analyze in R:

```{r}
library("phyloseq"); packageVersion("phyloseq")
library("DESeq2")
packageVersion("DESeq2")
library("ggplot2")

install.packages("RJSONIO")
install.packages("~/Documents/2017_Ecological_genomics_Pbio381/OTU analysis/biom_0.3.12.tar", repos = NULL, type = "source")
library("biom")



```


```{r}
x = read.table("~/Documents/2017_Ecological_genomics_Pbio381/OTU analysis/metagenome_predictions.L3.biom", sep = "\t", header = TRUE)
otumat = as(biom_data(x), "matrix")
OTU = otu_table(otumat, taxa_are_rows=TRUE)


mapping <- import_qiime_sample_data(mapfilename = 'R_map.txt')

phylo <- merge_phyloseq(OTU, mapping)
phylo

###############################################################################
###Test for DE KO terms between individuals that got sick and those that didn't
###############################################################################

final_pheno = phyloseq_to_deseq2(phylo, ~ Final_phenotype)
final_pheno_results = DESeq(final_pheno, test="Wald")
final_pheno_res = results(final_pheno_results)
summary(final_pheno_res)
head(final_pheno_res)

##build the table and output it to my computer:

alpha = 0.05
final_pheno_sigtab = final_pheno_res[which(final_pheno_res$padj < alpha), ]
final_pheno_sigtab= cbind(as(final_pheno_sigtab, "data.frame"), as(tax_table(phylo)[rownames(final_pheno_sigtab), ], "matrix"))
head(final_pheno_sigtab)
final_pheno_sigtab
write.table(final_pheno_sigtab, "Final_pheno_L3.txt", sep="\t")
```





